---
title: "GROUP-6 PROJECT"
author: "Lakshya Khorwal,Disha Monga,Shinjan Goswami,Priyansh Shrivastava,Yenni Akshya,Maddikuntla Pavan Kumar"
date: "`r Sys.Date()`"
always_allow_html: true
---

```{css, echo=FALSE}
.main-container {
  max-width: 100% !important;
  padding: 0 20px;}
pre.r {
  max-height: 250px;
  overflow-y: auto;
}
```

```{r setup, include=FALSE}
setwd("C:/ISI-D BSDS/DAURP/Project/Lakshya/Import-Export")
getwd()
s1=read.csv("Principal_Commodity_wise_export_for_the_year_202122.csv")
s2=read.csv("Principal_Commodity_wise_import_for_the_year_202122.csv")
s3=read.csv("Principal_Commodity_wise_export_for_the_year_202223.csv")
s4=read.csv("Principal_Commodity_wise_import_for_the_year_202223.csv")
setwd("C:/ISI-D BSDS/DAURP/Project/Lakshya/Import-Export/Disha")
library(readxl)
ss<-read_excel("G6.xlsx")
library(ggplot2)
library(plotly)
library(lattice)
library(dplyr)
library(corrplot)
library(scales)
library(tidyr)

```

## Introduction

Our data set is **'IMPORT-EXPORT of various commodities'**. The variables are: PRINCIPLE.COMMODITY, COUNTRY, UNIT, QUANTITY, Value(US$ millions)

The reason we chose this is because of the sheer importance of trade at a global level (as known to all by the very recent example of the halt in crude oil supply). We can observe the trend of a commodity, its demand, inflation, and the overall flow of trade.

Let's start off with how the data set looks like

##Export Value Heatmap


```{r heatmap_export-code, echo=FALSE, include=TRUE,message=FALSE, warning=FALSE, fig.width=12, fig.height=10, out.width="100%"}
# Note: echo=FALSE hides the R code so your slide only shows the beautiful graph!
s1.1=s1[-24050,]
p1=ggplot(s1.1, aes(x = COUNTRY, 
                               y = PRINCIPLE.COMMODITY, 
                               fill = Value.US..million.|>log(),
                               text = paste(
     "<b>Actual Value (US$M):", Value.US..million.)))
p1=p1+geom_tile(color = "white")+labs(title = "Expirt Value Heatmap",
        x = "Country", 
        y = "Commodity")+
   scale_fill_gradient(low = "darkblue",high = "yellow") +
   theme(axis.text.x = element_text(angle=90,size=7),
         axis.text.y = element_text(size = 6))
p1
p1_i<-ggplotly(p1)
p1_i
```

##Import Value heatmap

```{r heatmap_import-code, echo=FALSE, include=TRUE,message=FALSE, warning=FALSE, fig.width=12, fig.height=10, out.width="100%"}
# Note: echo=FALSE hides the R code so your slide only shows the beautiful graph!
s2.1=s2[-11928,]
p2=ggplot(s2.1, aes(x = COUNTRY, 
                               y = PRINCIPLE.COMMODITY, 
                               fill = Value.US..million.|>log(),
                               text = paste(
     "<b>Actual Value (US$M):", Value.US..million.)))
p2=p2+geom_tile(color = "white")+labs(title = "Import Value Heatmap",
        x = "Country", 
        y = "Commodity")+
   scale_fill_gradient(low = "darkblue",high = "yellow") +
   theme(axis.text.x = element_text(angle=90,size=7),
         axis.text.y = element_text(size = 6))
p2_i<-ggplotly(p2)
p2_i
```

One of the important factor is the valuation of dollar: 2022's $ is not the same as 2026's. As indicated by the source we chose: https://www.macrotrends.net/1329/us-dollar-index-historical-chart
here, the valuation of dollar did not change much during 2021-2022,2022-23

Because of this vastness of the data. We had decided to funnel down and use only 2 commodities over the two years.

simple results:
```{r simple, echo=FALSE, include=TRUE,message=FALSE, warning=FALSE, fig.width=12, fig.height=10, out.width="100%"}
trade_summary <- ss %>%
  group_by(`PRINCIPLE COMMODITY`, trade, Year) %>%
  summarize(
    Total_Quantity = sum(QUANTITY, na.rm = TRUE),
    Total_USD_Million = sum(`Value(US$ million)`, na.rm = TRUE),
    Top_Country = COUNTRY[which.max(QUANTITY)], 
    .groups = 'drop' 
  ) %>%
  arrange(`PRINCIPLE COMMODITY`, trade, Year) %>%
  as.data.frame()

trade_summary
```

another of the limitation is that there is no data set for 2020-21
hence the small sample.

```{css, echo=FALSE}
/* This forces the code blocks to have a vertical scrollbar if they get too long */
pre.r {
  max-height: 250px;
  overflow-y: auto;
}
```
```{r setup-clipboard, include=FALSE}
# This secretly activates the Copy button for the reader
library(xaringanExtra)
xaringanExtra::use_clipboard()
```
```{r jitter-plot, echo=TRUE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8, out.width="100%"}

ss$Year <- as.factor(ss$Year)
ss$Year <- as.numeric(as.character(ss$Year))
ss$`PRINCIPLE COMMODITY` <- as.factor(ss$`PRINCIPLE COMMODITY`)
ss$trade <- as.factor(ss$trade)
ss_avg <- tapply(ss$`Value(US$ million)`, 
                 list(ss$Year, ss$`PRINCIPLE COMMODITY`, ss$trade), 
                 mean, na.rm = TRUE) |>
  as.table() |>
  as.data.frame() |>
  # Make sure the name here perfectly matches your column name!
  setNames(c("Year", "PRINCIPLE COMMODITY", "trade", "mean_value")) |>
  subset(!is.na(mean_value))

ss_avg <- ss %>%
  group_by(Year, `PRINCIPLE COMMODITY`, trade) %>%
  summarize(mean_value = mean(`Value(US$ million)`, na.rm = TRUE), .groups = 'drop')

d <- ggplot() +
  geom_jitter(
    data = ss, 
    aes(
      x = as.character(Year), 
      y = `Value(US$ million)`, 
      color = trade,   # Color-code by import vs export
      shape = trade,   # Different shapes for import vs export
      text = paste(
        "Country:", COUNTRY,
        "<br>Trade Type:", trade,
        "<br>Commodity:", `PRINCIPLE COMMODITY`,
        "<br>Value (US$M):", `Value(US$ million)`
        )),
    width = 0.1, alpha = 0.5, size = 2) +
  
  # LAYER B: The line connecting the averages
  geom_line(
    data = ss_avg,
    aes(
      x = as.character(Year), 
      y = mean_value, 
      color = trade, 
      group = trade,
      text = paste(
        "<b>AVERAGE", toupper(trade), "</b>",
        "<br>Commodity:", `PRINCIPLE COMMODITY`,
        "<br>Mean Value (US$M):", round(mean_value, 2)
      )
    ),
    linewidth = 1.2
  ) +
  
  # LAYER C: Solid black points on the average line
  geom_point(
    data = ss_avg,
    aes(
      x = as.character(Year), 
      y = mean_value, 
      shape = trade, 
      text = paste(
        "<b>AVERAGE", toupper(trade), "</b>",
        "<br>Commodity:", `PRINCIPLE COMMODITY`,
        "<br>Mean Value (US$M):", round(mean_value, 2)
      )
    ),
    size = 3, color = "black" 
  ) +
  
  # Split the graph into two panels for Silver and Iron Ore
  facet_wrap(~ `PRINCIPLE COMMODITY`) +                 
  
  # Use a pseudo-log scale so massive outliers (like China) don't crush the smaller countries
  scale_y_continuous(trans = pseudo_log_trans(base = 10)) + 
  theme_minimal() +
  labs(
    title = "Interactive Trade Value by Year",
    x = "Year",
    y = "Value (US$ Million) - Log Scaled",
    color = "Trade Type",
    shape = "Trade Type"
  ) +
  theme(legend.position = "bottom")
d_i<- ggplotly(d, tooltip="text")
d_i
```

Another way to represent this graph
```{r cloud_3d, , echo=TRUE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8, out.width="100%"}
ss$Year <- as.factor(ss$Year)
ss$`PRINCIPLE COMMODITY` <- as.factor(ss$`PRINCIPLE COMMODITY`)
ss$Year <- as.character(ss$Year)
ss$`PRINCIPLE COMMODITY` <- as.character(ss$`PRINCIPLE COMMODITY`)
# Build the interactive 3D plot
fig <- plot_ly(
  data = ss, 
  x = ~as.character(Year), 
  y = ~`PRINCIPLE COMMODITY`, 
  z = ~`Value(US$ million)`|>log(), 
  color = ~trade, # Color the dots by import/export
  
  # Custom hover text
  text = ~paste(
    "Country:", COUNTRY,
    "<br>Trade Type:", trade,
    "<br>Value (US$M):", `Value(US$ million)`
  ),
  hoverinfo = "text",
  
  # Define the plot type
  type = "scatter3d", 
  mode = "markers",
  marker = list(size = 5, opacity = 0.7)
)

# Add titles to the 3D axes
fig <- fig %>% layout(
  title = "Interactive 3D Trade Value",
  scene = list(
    xaxis = list(title = "Year", type = "category"),
    yaxis = list(title = "Commodity"),
    zaxis = list(title = "Value (US$ M)")))
fig

```

The trend in each country
```{css, echo=FALSE}
/* This forces the code blocks to have a vertical scrollbar if they get too long */
pre.r {
  max-height: 250px;
  overflow-y: auto;
}
```
```{r xyplot, echo=TRUE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8, out.width="100%"}

e <- ggplot(ss, aes(
  x = as.character(Year), 
  y = `Value(US$ million)`, 
  group = interaction(COUNTRY, trade),
  color = trade,
  
  # Build the custom hover tooltip
  text = paste(
    "Country:", COUNTRY,
    "<br>Trade Type:", trade,
    "<br>Value (US$M):", `Value(US$ million)`
  )
)) +
  geom_line(alpha = 0.4) + 
  geom_point(size = 1.5, alpha = 0.6) + 
  
  facet_wrap(~ `PRINCIPLE COMMODITY`) + 
  scale_y_continuous(trans = pseudo_log_trans(base = 10)) + 
  theme_minimal() +
  labs(
    title = "Interactive Trade Value Trajectories by Country",
    x = "Year",
    y = "Value (US$ Million)"
  )

e_i<-ggplotly(e, tooltip = "text")
e_i
```


corelation plot
```{css, echo=FALSE}
/* This forces the code blocks to have a vertical scrollbar if they get too long */
pre.r {
  max-height: 8em;
  overflow-y: auto;
}
```
```{r corrplot, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=8, out.width="33%", fig.align="center", fig.show="hold"}
ss$Metric <- paste(ss$`PRINCIPLE COMMODITY`, ss$trade, sep = "_")
ss_2021 <- subset(ss, Year == "2021")
wide_matrix_1 <- xtabs(`Value(US$ million)` ~ COUNTRY + Metric,
                     data = ss_2021)
cor_matrix <- cor(as.matrix(wide_matrix_1))
corrplot(cor_matrix, 
         method = "shade",       # Fills the grid with color (blue = positive, red = negative)
         type = "upper",         # Hides the bottom half since it's a mirrored reflection
         addCoef.col = "black",  # Overlays the exact correlation numbers in black text
         tl.col = "black",       # Makes the text labels black (default is sometimes red)
         tl.srt = 90,            # Rotates the top text 45 degrees so it fits nicely
         diag = FALSE,           # Hides the diagonal line of 1.0s (a variable is always 100% correlated with itself)
         title = "2021 Correlation", mar=c(0,0,2,0))


ss$Metric <- paste(ss$`PRINCIPLE COMMODITY`, ss$trade, sep = "_")
ss_2022 <- subset(ss, Year == "2022")
wide_matrix_2 <- xtabs(`Value(US$ million)` ~ COUNTRY + Metric,
                     data = ss_2022)
cor_matrix <- cor(as.matrix(wide_matrix_2))
corrplot(cor_matrix, 
         method = "shade",       # Fills the grid with color (blue = positive, red = negative)
         type = "upper",         # Hides the bottom half since it's a mirrored reflection
         addCoef.col = "black",  # Overlays the exact correlation numbers in black text
         tl.col = "black",       # Makes the text labels black (default is sometimes red)
         tl.srt = 90,            # Rotates the top text 45 degrees so it fits nicely
         diag = FALSE,           # Hides the diagonal line of 1.0s (a variable is always 100% correlated with itself)
         title = "2022 Correlation", mar=c(0,0,2,0))
```


Inflation

inflation formula: $\frac{\text{Current Year} - \text{Past Year}}{\text{Past Year}} \times 100$.
```{css, echo=FALSE}
/* This forces the code blocks to have a vertical scrollbar if they get too long */
pre.r {
  max-height: 8em;
  overflow-y: auto;
}
```
```{r inflation, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=8, out.width="33%" }
inflation_summary <- ss %>%
  # Group by Commodity, Trade Type, and Year
  group_by(`PRINCIPLE COMMODITY`, trade, Year) %>%
  
  # Calculate the true average unit price for that year
  summarize(
    Total_Value = sum(`Value(US$ million)`, na.rm = TRUE),
    Total_Qty = sum(QUANTITY, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    Avg_Yearly_Price = ifelse(Total_Qty > 0, Total_Value / Total_Qty, NA)
  ) %>%
  
  pivot_wider(
    id_cols = c(`PRINCIPLE COMMODITY`, trade),
    names_from = Year,
    values_from = Avg_Yearly_Price,
    names_prefix = "Price_"
  ) %>%
  
  mutate(
    Inflation_Percent = ((Price_2022 - Price_2021) / Price_2021) * 100
  ) %>%
  arrange(desc(Inflation_Percent))

inflation_summary
```

Cumulative
```{r cumulative,echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5, out.width="33%", , fig.align="center", fig.show="hold"}
#Iron ore
##import 2021
iron_import_2021 <- ss %>%
  filter(`PRINCIPLE COMMODITY` == "IRON ORE", trade == "import", Year == 2021) %>%
  group_by(COUNTRY) %>%
  summarise(total_value = sum(`Value(US$ million)`, na.rm = TRUE)) %>%
  arrange(desc(total_value)) %>%
  mutate(
    cum_value = cumsum(total_value),
    cum_percent = cum_value / sum(total_value) * 100)
iron_import_2021$cum_percent
iron_import_2021$COUNTRY[1:10]
cu_1<-ggplot(iron_import_2021, aes(x = reorder(COUNTRY, -total_value))) +
  geom_bar(aes(y = total_value), stat = "identity") +
  geom_line(aes(y = cum_percent * max(total_value) / 100, group = 1),
            color="green") +
  geom_point(aes(y = cum_percent * max(total_value) / 100), color="darkgreen", size = 2) +
  coord_flip() +
  ggtitle("Iron Ore Import 2021")
ggplotly(cu_1)

##import 2022
iron_import_2022 <- ss %>%
  filter(`PRINCIPLE COMMODITY` == "IRON ORE", trade == "import", Year == 2022) %>%
  group_by(COUNTRY) %>%
  summarise(total_value = sum(`Value(US$ million)`, na.rm = TRUE)) %>%
  arrange(desc(total_value)) %>%
  mutate(
    cum_value = cumsum(total_value),
    cum_percent = cum_value / sum(total_value) * 100)
iron_import_2022$cum_percent
iron_import_2022$COUNTRY[1]
iron_import_2022$COUNTRY[1:10]
cu_2<-ggplot(iron_import_2022, aes(x = reorder(COUNTRY, -total_value))) +
  geom_bar(aes(y = total_value), stat = "identity") +
  geom_line(aes(y = cum_percent * max(total_value) / 100, group = 1),
            color="green") +
  geom_point(aes(y = cum_percent * max(total_value) / 100), color="darkgreen", size = 2) +
  coord_flip() +
  ggtitle("Iron Ore Import 2022")
ggplotly(cu_2)
##export2021
iron_export_2021 <- ss %>%
  filter(`PRINCIPLE COMMODITY` == "IRON ORE", trade == "export", Year == 2021) %>%
  group_by(COUNTRY) %>%
  summarise(total_value = sum(`Value(US$ million)`, na.rm = TRUE)) %>%
  arrange(desc(total_value)) %>%
  mutate(
    cum_value = cumsum(total_value),
    cum_percent = cum_value / sum(total_value) * 100)
iron_export_2021$cum_percent
iron_export_2021$COUNTRY[1:10]
cu_3<-ggplot(iron_export_2021, aes(x = reorder(COUNTRY, -total_value))) +
  geom_bar(aes(y = total_value), stat = "identity") +
  geom_line(aes(y = cum_percent * max(total_value) / 100, group = 1),
            color="green") +
  geom_point(aes(y = cum_percent * max(total_value) / 100), color="darkgreen", size = 2) +
  coord_flip() +
  ggtitle("Iron Ore Export 2021")
ggplotly(cu_3)

##export2022
iron_export_2022 <- ss %>%
  filter(`PRINCIPLE COMMODITY` == "IRON ORE", trade == "export", Year == 2022) %>%
  group_by(COUNTRY) %>%
  summarise(total_value = sum(`Value(US$ million)`, na.rm = TRUE)) %>%
  arrange(desc(total_value)) %>%
  mutate(
    cum_value = cumsum(total_value),
    cum_percent = cum_value / sum(total_value) * 100)
iron_export_2022$cum_percent
iron_export_2022$COUNTRY[1:10]
cu_4<-ggplot(iron_export_2022, aes(x = reorder(COUNTRY, -total_value))) +
  geom_bar(aes(y = total_value), stat = "identity") +
  geom_line(aes(y = cum_percent * max(total_value) / 100, group = 1),
            color="green") +
  geom_point(aes(y = cum_percent * max(total_value) / 100), color="darkgreen", size = 2) +
  coord_flip() +
  ggtitle("Iron Ore Export 2022")
ggplotly(cu_4)
```
